Topological approach to finding blackholes in directed networks

### 2019-12-03 blackholes/pavt/article.tex

According to the definitions above if some node is not a root of a blackhole, it can be omitted, as set of roots uniquely identify a blackhole.
Every non-root node in a blackhole is reachable from one of the roots. Therefore, if we know reachability matrix for a given graph, we can skip bloated non-basis combinations of nodes.
Of course, we could straightforward build the reachability matrix. But it takes $O(V^3)$ time. \FIXME{определить $G=(V, E)$} Even though we decreased graph size in preprocessing step, it still can result in lots of computation. That is why we will only try to partially skip duplicates and the rest will be filtered later.

Let's consider a closure of a single blackhole root. Let this closure to be acquired by depth-first search, which basically means, that
we can iterate over these nodes in topologically sorted order. Also, while traversing this graph, we can calculate size of closure for each node
in the given roots closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
conclude, that every node with index less than $i$ in the topological sort will be reachable from $v[i]$. \FIXME{определить v[i]}

%
\section{Experimental Results}
To demonstrate efficiency of our approach, we conducted a series of experiments. We ran the both algorithms (New Jersey and TopSort) on a 
set of graphs.

@@

\begin{table}[]
\caption{Experimental results}
\label{tabular:tableresults}
\begin{center}
\begin{tabular}{c|c|c|c|c|c|c|c|c}
Graph & Type & Algorithm & Blackholes & Time & Algorithm & Blackholes & Time \\
\hline
RMAT & 04 & NewJersey & 1 & 0.000265422 & TopSort & 1 & 0.000115912 \\
RMAT & 06 & NewJersey & 1 & 0.0121347 & TopSort & 1 & 0.00034812 \\
RMAT & 08 & NewJersey & 1 & 1.66685 & TopSort & 1 & 0.00146486 \\
RMAT & 10 & NewJersey & 6 & 1200 & TopSort & 10 & 0.00684045 \\
RMAT & 12 & NewJersey & 29 & 1200 & TopSort & 971738 & 1200 \\
RMAT & 14 & NewJersey & 187 & 1200 & TopSort & 735128 & 1200 \\
RMAT & 16 & NewJersey & 1192 & 1200 & TopSort & 373705 & 1200 \\
RMAT & 18 & NewJersey & 5176 & 1200 & TopSort & 10286 & 1200 \\
RMAT & 20 & NewJersey & 617 & 1200 & TopSort & 48867 & 1200 \\
RMAT & 22 & NewJersey & 98 & 1200 & TopSort & 0 & 1200 \\
SSCA2 & 04 & NewJersey & 16 & 0.154763 & TopSort & 16 & 0.000189925 \\
SSCA2 & 06 & NewJersey & 11 & 1200 & TopSort & 70 & 0.00121528 \\
SSCA2 & 08 & NewJersey & 42 & 1200 & TopSort & 263314 & 1200 \\
SSCA2 & 10 & NewJersey & 85 & 1200 & TopSort & 114411 & 1200 \\
SSCA2 & 12 & NewJersey & 343 & 1200 & TopSort & 5870 & 1200 \\
SSCA2 & 14 & NewJersey & 1433 & 1200 & TopSort & 323671 & 1200 \\
SSCA2 & 16 & NewJersey & 5829 & 1200 & TopSort & 893376 & 1200 \\
SSCA2 & 18 & NewJersey & 23179 & 1200 & TopSort & 224567 & 1200 \\
SSCA2 & 20 & NewJersey & 92347 & 1200 & TopSort & 24199 & 1200 \\
SSCA2 & 22 & NewJersey & 370770 & 1200 & TopSort & 0 & 1200 \\
UR & 04 & NewJersey & 1 & 0.000223768 & TopSort & 1 & 0.000115029 \\
UR & 06 & NewJersey & 1 & 0.0170713 & TopSort & 1 & 0.000331235 \\
UR & 08 & NewJersey & 1 & 2.30256 & TopSort & 1 & 0.00147483 \\
UR & 10 & NewJersey & 1 & 240.822 & TopSort & 1 & 0.00670724 \\
UR & 12 & NewJersey & 0 & 1200 & TopSort & 1 & 0.034011 \\
UR & 14 & NewJersey & 0 & 1200 & TopSort & 1 & 0.47257 \\
UR & 16 & NewJersey & 0 & 1200 & TopSort & 1 & 3.05003 \\
UR & 18 & NewJersey & 0 & 1200 & TopSort & 1 & 14.5788 \\
UR & 20 & NewJersey & 0 & 1200 & TopSort & 1 & 67.7176 \\
UR & 22 & NewJersey & 0 & 1200 & TopSort & 1 & 373.244
\end{tabular}
\end{center}
\end{table}

>>>>

According to the definitions above if some node is not a root of a blackhole, it can be omitted, as set of roots uniquely identify a blackhole.
Every non-root node in a blackhole is reachable from one of the roots. Therefore, if we know reachability matrix for a given graph, we can skip bloated non-basis combinations of nodes.

\FIXME{определить $G=(V, E)$}
\FIXME{определить v[i]}

Of course, we could straightforward build the reachability matrix. But, it takes $O(V^3)$ time. Even though we decreased graph size
in preprocessing step, it still can result in lots of computation. It's worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
Difficult precalculation will affect the time of first blackhole discovery. That is why, we will only try to partially skip duplicates and the rest will
be filtered later in lazy manner.

Let's consider a closure of a single blackhole root. Let this closure to be acquired by depth-first search, which basically means, that
we can iterate over these nodes in topologically sorted order. Also, while traversing this graph, we can calculate size of closure for each node
in the given roots closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
conclude, that every node with index less than $i$ in the topological sort will be reachable from $v[i]$. 
TopSortFilter algorithm:
\begin{enumerate}
    \item For every node 
    \item Calculate topological sort order
    \item Calculate closure size for every vertex
    \item For every node
\end{enumerate}
%
\section{Experimental Results}
To demonstrate efficiency of our approach, we conducted a series of experiments. We ran the both algorithms (New Jersey and TopSort) on a 
set of graphs.

iBlackhole algorithm was implemented in accordance with the pseudocode in the original article. Divide and Conquer heuristic were not used for it.
Our algorithm first applies condensation-based preprocessing and then uses BruteForce with partial reachability heuristic implemented via topsort.
Both algorithms filter duplicates

@@

Please, mention, that "blackhole found" means that we have a set of vertex numbers in memory ready to be printed out and it's already validated.

\begin{table}[]
\caption{Experimental results}
\label{tabular:tableresults}
\begin{center}
\begin{tabular}{c|c|c|c|c|c|c|c|c}
Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
\hline
RMAT & 04 & NewJersey & 1 & 0.000265422 & TopSort & 1 & 0.000115912 \\
RMAT & 06 & NewJersey & 1 & 0.0121347 & TopSort & 1 & 0.00034812 \\
RMAT & 08 & NewJersey & 1 & 1.66685 & TopSort & 1 & 0.00146486 \\
RMAT & 10 & NewJersey & 6 & 1200 & TopSort & 10 & 0.00684045 \\
RMAT & 12 & NewJersey & 29 & 1200 & TopSort & 971738 & 1200 \\
RMAT & 14 & NewJersey & 187 & 1200 & TopSort & 735128 & 1200 \\
RMAT & 16 & NewJersey & 1192 & 1200 & TopSort & 373705 & 1200 \\
RMAT & 18 & NewJersey & 5176 & 1200 & TopSort & 10286 & 1200 \\
RMAT & 20 & NewJersey & 617 & 1200 & TopSort & 48867 & 1200 \\
RMAT & 22 & NewJersey & 98 & 1200 & TopSort & 0 & 1200 \\
SSCA2 & 04 & NewJersey & 16 & 0.154763 & TopSort & 16 & 0.000189925 \\
SSCA2 & 06 & NewJersey & 11 & 1200 & TopSort & 70 & 0.00121528 \\
SSCA2 & 08 & NewJersey & 42 & 1200 & TopSort & 263314 & 1200 \\
SSCA2 & 10 & NewJersey & 85 & 1200 & TopSort & 114411 & 1200 \\
SSCA2 & 12 & NewJersey & 343 & 1200 & TopSort & 5870 & 1200 \\
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Experimental results. Part 2}
\label{tabular:tableresults2}
\begin{center}
\begin{tabular}{c|c|c|c|c|c|c|c|c}
Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
SSCA2 & 14 & NewJersey & 1433 & 1200 & TopSort & 323671 & 1200 \\
SSCA2 & 16 & NewJersey & 5829 & 1200 & TopSort & 893376 & 1200 \\
SSCA2 & 18 & NewJersey & 23179 & 1200 & TopSort & 224567 & 1200 \\
SSCA2 & 20 & NewJersey & 92347 & 1200 & TopSort & 24199 & 1200 \\
SSCA2 & 22 & NewJersey & 370770 & 1200 & TopSort & 0 & 1200 \\
UR & 04 & NewJersey & 1 & 0.000223768 & TopSort & 1 & 0.000115029 \\
UR & 06 & NewJersey & 1 & 0.0170713 & TopSort & 1 & 0.000331235 \\
UR & 08 & NewJersey & 1 & 2.30256 & TopSort & 1 & 0.00147483 \\
UR & 10 & NewJersey & 1 & 240.822 & TopSort & 1 & 0.00670724 \\
UR & 12 & NewJersey & 0 & 1200 & TopSort & 1 & 0.034011 \\
UR & 14 & NewJersey & 0 & 1200 & TopSort & 1 & 0.47257 \\
UR & 16 & NewJersey & 0 & 1200 & TopSort & 1 & 3.05003 \\
UR & 18 & NewJersey & 0 & 1200 & TopSort & 1 & 14.5788 \\
UR & 20 & NewJersey & 0 & 1200 & TopSort & 1 & 67.7176 \\
UR & 22 & NewJersey & 0 & 1200 & TopSort & 1 & 373.244
\end{tabular}
\end{center}
\end{table}

\pagebreak
%
